# r-consortium-proposal
Application for https://www.r-consortium.org/projects/call-for-proposals

```{r, echo=FALSE, message=FALSE}
library("knitr")
library("googlesheets")
```



# R Consortium Proposal: School of Data Material Development

[Heidi Seibold](http://www.ebpi.uzh.ch/en/aboutus/departments/biostatistics/teambiostats/seibold.html),
[Camila Salazar](),
[David Opoku](),
[Cedric Lombion]()

## The Problem
R has great tools for accessing and analysing open data sources, which would be
hugely beneficial to journalists.  However there is a low level of awareness of
R in the journalism community and specialized courses -- especially in
languages other than English -- are not available.

Journalists can profit from data skills because they allow them to access
diverse data, do more profound research and communicate through visualisations.

In order to be
able to tailor material for the needs of journalists we sent out a survey to
journalists all over the world to assess what those needs are. From the survey
we learned that ... **TODO** ...

```{r, echo=FALSE, message=FALSE}
responses_sheet <- gs_title("R course for journalists  (Responses)")
responses <- gs_read(responses_sheet)

tools <- responses["What tools do you normally use to work with data? (select all that apply)"]
```



## The Plan

#### Target audience
Journalists

#### Idea
We aim to create introductory content about R which lowers the barrier to entry 
for journalists by using examples, datasets and tips that are tailored to the 
needs and workflow of journalists using R as part of their work.
There exist many introductory R courses for a broad audience and we are planning
to build ontop of them, i.e., do the basic introduction with available material
and and this use specialised
material for journalists, which still needs to be developed. 

#### Goals

- Write introductory learning content about R tailored for journalists
- Design a content format that encourages journalists to learn without feeling overwhelmed
- Select and organise existing learning resources about R to direct journalists to them
- Provide all in four languages: English, French, Spanish and German

#### Deliverables

- x modules about learning R for journalists, from "why would I want to use R?" to 
  "using R in a newsroom workflow"
- A website which presents this content in an attractive format
- Learning tracks which guide journalists through additional resources after 
  they have finished the introduction
- Make all the material available via GitHub to use it and develop it further.


#### Timeline
**TODO: timeline**


#### Likely failure modes and how to recover
As we will prepare the material in a modular way, it will not be a big problem
if we do not manage preparing as much material as we now anticipate. In this
case we will just have less modules, but quality of those will be ensured.



## How Can The ISC Help?

```{r, echo=FALSE, message=FALSE}

costs_sheet <- gs_title("ScoDa R consortium proposal: Costs")
costs <- gs_read(costs_sheet)
costs$Notes[is.na(costs$Notes)] <- ""

material <- subset(costs, Info == "material", 
                   select = c("Work", "Funding requested", 
                              "School of Data co-funding"))
workshop <- subset(costs, Info == "workshop",  
                   select = c("Work", "Funding requested", 
                              "School of Data co-funding"))
```

We would like to request financial support of `r sum(material["Funding requested"])` 
USD from the ISC to develop material for journalists who would like to learn R. 
The School of Data can support the material development in guidance for the 
material writing as well as in communication and maintainance.
The development costs can be split up as follows: 
```{r, echo=FALSE}
kable(material)
```


If additional support is possible, we would like to organize workshops in Ghana, 
Costa Rica, France and Switzerland. The financial support needed are 
`r sum(workshop["Funding requested"])` USD and can be split up as follows:
```{r, echo=FALSE}
kable(workshop)
```


## Dissemination
The School of Data has a broad community with many members working in journalism.
We will promote the material among the community and beyond, by sending out
newsletters, [tweets](https://twitter.com/SchoolOfData) and informations on the main 
[School of Data website](https://schoolofdata.org/) as well
as the websites of the chapters in the different countries. We will publish
the materials under CC-BY license and make it easy for others to build upon the
content and extend it.





