---
title: "Skills Lesson: Skills Lesson Name"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library("learnr")
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This is where a very short introduction goes.

Possible skills lessons could be:

- Skills Lesson: Reading Data into R
- Skills Lesson: Data Preparation
- Skills Lesson: Visualising Data

Skills lessons should be very short and simple.

- Show important packages with 1 or 2 simple examples. 
- Link to other pages for *further reading*.

### Naming conventions
Recipe files should be named `language-skills-name.Rmd` (example `de-skills-reading-data.Rmd`)

- `language` should be `en` for English, `de` for German, `es` for Spanish and `fr` for French. File names should all be in English in order to know which lessons belong together.
- `skills` is the same word for all skills lessons (look at the recipe template for recipe naming conventions).
- `name` should be a descriptive name (e.g. `visualisation` for a data visualisation skills lesson).

Titles of skills lessons should be `Skills Lesson: Skills Lesson Name` with upper case letters in English, 
`Kompetenzlektion: Name` in German, ??? in Spanish and ??? in French.


## Important packages and functions

For this skill packages `ggplot2` and `dplyr` are especially important. 
With them you can do things like
```{r plot1, exercise = TRUE}
library("ggplot2")
d <- ggplot(diamonds, aes(carat, price))
d + geom_point(alpha = 1/10)
```
and
```{r summarise, exercise = TRUE, message=FALSE}
library("dplyr")
iris %>%
  group_by(Species) %>%
  summarise_all(mean)
```


## Further reading

- [RStudio cheatsheets](https://www.rstudio.com/resources/cheatsheets/)
on **this** and **that**.
- The [dplyr website](http://dplyr.tidyverse.org/).



