---
title: Recipe for analysis of the dual-use goods exports
author: "Heidi Seibold and Peter B. Pearman"
date: "April 25, 2017"
output: learnr::tutorial
runtime: shiny_prerendered
---


```{r  echo=FALSE}
# set the knitr option so that no chunks produce messages 
# silently load the learnr library so that the file will run as a shiny document
knitr::opts_chunk$set(message = FALSE)
library("learnr")
```

```{r}

```

# **Analysis of Dual-Use Goods Exports**

### Authors: Heidi Seibold and Peter B. Pearman



## What are dual-use goods?
Dual-use goods are goods that can be used for civil and military purposes. 
In Switzerland, these goods are controlled by special legislation,
unlike in other countries where they are looked at as conventional arms exports. 

The state secretariat for economic affairs (SECO) in Switzerland decides which
dual-use goods may be exported and also 
[provides the data](https://www.seco.admin.ch/seco/de/home/Aussenwirtschaftspolitik_Wirtschaftliche_Zusammenarbeit/Wirtschaftsbeziehungen/exportkontrollen-und-sanktionen/industrieprodukte--dual-use--und-besondere-militaerische-gueter/statistik/2015.html) 
on a quarterly basis.

The [SRF data](srf.ch/data) team - the data journalism team of Swiss Radio and TV -
has been repeatedly analysing these data in the past years. 

The reasons for using R for this task are various:

- R can read in Excel data (SECO provides the data in .xlsx format).
- R allows for automation of data cleaning and analyses that have to be repeated
- R can export data to .csv so it can be used in other programs (e.g. Excel).
- R produces good graphics and allows for individualised graphical themes.
- Reproducible documents (e.g. HTML or PDF) can be created with R, while changes  and updates can be included without major effort.
- R allows you to document how you have transformed the data, allowing accountablility and providing readers with tranparency. 
- With R, reproducible documents (e.g. HTML or PDF) can be created, changes are 
included without major efforts.

You can find the entire analysis on the [SRF data GitHub repository](https://github.com/srfdata/2017-01-dual-use)
and the article from January 2017 on the 
[SRF website](http://www.srf.ch/news/schweiz/schweiz-bewilligt-millionendeal-mit-dem-irak).

So let's jump right into the School of Data analysis pipeline!  As a reminder, you can find it [here](https://schoolofdata.org/methodology/).

## Define

At the beginning of 2017 SRF data analyst Timo Grossenbacher wanted to answer the 
question   
**Which dual-use goods were exported to which countries in each quarter of the 
years 2012 to 2016?**

Having cleaned and visualised the data he realised that suddenly in 2016 Switzerland started exporting dual-use goods to Irak after not having done so in the years prior. This became the story in the January 2017 [article](http://www.srf.ch/news/schweiz/schweiz-bewilligt-millionendeal-mit-dem-irak) .

We want to answer the same question and in order to have a clean project, please create
a new folder on your computer in which you will store all the details of this project.
We propose the following:

- Create a folder called `switzerland-dual-use`
- Within this folder create two additional folders called `data_raw` and `data_clean`

```{}
└── switzerland-dual-use
    ├── data_clean
    └── data_raw
```


## Find 

The data can be found (in German) on the [SECO website](https://www.seco.admin.ch/seco/de/home/Aussenwirtschaftspolitik_Wirtschaftliche_Zusammenarbeit/Wirtschaftsbeziehungen/exportkontrollen-und-sanktionen/industrieprodukte--dual-use--und-besondere-militaerische-gueter/statistik/2015.html). SRF also keeps the data available in their [GitHub repo](https://github.com/srfdata/2017-01-dual-use/tree/master/analysis/input).



## Get

The data can be downloaded by clicking on the data links on the 
[SECO website](https://www.seco.admin.ch/seco/de/home/Aussenwirtschaftspolitik_Wirtschaftliche_Zusammenarbeit/Wirtschaftsbeziehungen/exportkontrollen-und-sanktionen/industrieprodukte--dual-use--und-besondere-militaerische-gueter/statistik/2015.html). Get the data files and store them in the folder `data_raw`.

Your file structure should now look like this,
```{}
└── switzerland-dual-use
    ├── data_clean
    └── data_raw
        ├── 1_Quartal_2016_Bewilligungen.xlsx
        ├── 2. Quartal 2016 Erteilte Ausfuhrbewilligungen.xlsx
        ├── 3. Quartal 2016 Erteilte Ausfuhrbewilligungen.xlsx
        ├── 4. Quartal 2016 Erteilte Ausfuhrbewilligungen.xlsx
        ├── Stat_IP_2015_Bewilligungen.xlsx
        ├── Statistik Dual-Use 2014 Elic.xlsx
        ├── Statistik Dual-Use Jan. 2012 - Sep. 2014 Tracker.xlsx
        └── Statistik ML 2005-2013.xlsx
```
i.e. you should have the dual-use Excel files corresponding to years 2012 to 2016.


If you have already worked through the Introduction lesson, you probably already have RStudio installed and can try running the following code in the Console window.  If you haven't worked the Introduction lesson, you can either do that now, or just continue reading through this tutorial and do the exercises when you come to them.

Let's try reading in the data from first two quarters of 2016:

1. Open your R editor (e.g. RStudio) 
2. Set the R working directory to your `switzerland-dual-use` directory. 
To do this, first enter 
```{r, eval=FALSE}
getwd()
```
in the console. This will show your current working directory.
E.g. for me it shows `[1] "/home/heidi/Documents"`. Next use `setwd()` to change 
to the correct diretory. For me it is `/home/heidi/Documents/switzerland-dual-use` 
and so I have to type
```{r, eval=FALSE}
setwd("switzerland-dual-use")
```
or equivalently 
```{r, eval=FALSE}
setwd("/home/heidi/Documents/switzerland-dual-use")
```
or assuming you have this file in the directory `switzerland-dual-use` and its file structure is as described above, you can use RStudio's menu. Click on `Session`, then on `Set Working Directory` and manually specify the path including `switzerland-dual-use`.

### SET YOUR R WORKING DIRECTORY NOW 

To read Excel data into R, we need a special R package called
`readxl`. To install the package run the following line in the Console window of RStudio:
```{r, eval=FALSE}
install.packages("readxl")
```
To load the package, so you can access the functions, run the following in the Console window:
```{r}
library("readxl") 
```
Now we can read in the data from the four quarters of 2016:
```{r eval=FALSE}
elic_2016_q1_raw <- read_excel(path = "data_raw/1_Quartal_2016_Bewilligungen.xlsx")
elic_2016_q2_raw <- read_excel(path = "data_raw/2. Quartal 2016 Erteilte Ausfuhrbewilligungen.xlsx")
elic_2016_q3_raw <- read_excel(path = "data_raw/3. Quartal 2016 Erteilte Ausfuhrbewilligungen.xlsx")
elic_2016_q4_raw <- read_excel(path = "data_raw/4. Quartal 2016 Erteilte Ausfuhrbewilligungen.xlsx")
```
```{r echo=FALSE}
library(ddj)
data(list=c("elic_2016_q1_raw","elic_2016_q2_raw","elic_2016_q3_raw","elic_2016_q4_raw"))
```


## Verify
To check whether reading in the files worked and the data are as we wanted, there are a couple of useful commands in R. Let's now first work with `elic_2016_q1_raw`:

- `names()` shows the column names of the data frame
```{r}
names(elic_2016_q1_raw)
```
- `head()` shows the first few rows of the data frame
```{r}
head(elic_2016_q1_raw)
```
- `tail()` shows the last few rows of the data frame
```{r}
tail(elic_2016_q1_raw)
```
- `summary()` shows a summary of the data. For numeric variables such as the price (`Wert [CHF]`) this function 
shows the [five number summary](https://en.wikipedia.org/wiki/Five-number_summary) and the mean (average).  For character variables such as receiving country (`Bestimmungsland`) the function shows the variable's length and informs that it is of the character class and mode.
```{r}
summary(elic_2016_q1_raw)
```
- `str()` shows the type of each variable (here `num` for numeric and `chr` for character) and the
first few values
```{r}
str(elic_2016_q1_raw)
```


## Clean

In this section we will need the `dplyr` package
```{r}
library("dplyr")   ## if not available install by typing install.packages("dplyr")
```


As a first step we want to combine the four 2016 data sets. You might have already
realised that the first data set has one collumn that is just `NA`. `NA` is the 
missing value indicator in R. Thus, this collumn does not contain any 
information. Let's see what happens to this collumn if we combine the four 
data sets using function `bind_rows()`.  `bind_rows()` keeps all collumns in all data sets and fills a column with `NA` when a data set does not contain the column. Also it created a new collumn called`Quartal`. It did that because we told it that the `.id` collumn is `Quartal` and the values for each `Quartal` are given in the quotation marks `"`.

```{r}
names(elic_2016_q1_raw)
names(elic_2016_q2_raw)

elic_2016_1 <- bind_rows("16/01" = elic_2016_q1_raw, 
                         "16/02" = elic_2016_q2_raw, 
                         "16/03" = elic_2016_q3_raw, 
                         "16/04" = elic_2016_q4_raw, 
                         .id = "Quartal")

names(elic_2016_1)
str(elic_2016_1)
```

```{r prepare-clean1, echo= FALSE}
library(ddj)
data("elic_2016_1")
#load(url("https://github.com/pbpearman/r-consortium-proposal/raw/master/material/lessons/switzerland-dual-use/data_clean/elic_2016_1.RData"))
#load(url("https://github.com/school-of-data/r-consortium-proposal/raw/master/r-package/data/elic_2016_1.RData"))

```

### Exercise 1
Use head() with the `n=` option to get the first 10 rows of elic_2016_1 
```{r check_values, exercise=TRUE, exercise.setup = 'prepare-clean1', cache=FALSE}

```
```{r check_values-solution}
head(elic_2016_1, n=10)
```




Now we do not want to keep the collumn with just `NA` values and we want to 
rename a few variables. We can do this with function `select()`:

- To just select a variable, name it (e.g. `Quartal`)
- To select and rename a variable, put `newname = oldname` (e.g. `GN = Geschäftsnummer`)
- For variables with spaces or special symbols, use a single back-quote to show where it starts and ends
  (see e.g. `Wert [CHF]`)
- To select all except one variable, use the minus symbol 
  (e.g. `select(.data = elic_2016_3, -Art)` below)
```{r}
elic_2016_2 <- select(.data = elic_2016_1,
                      GN = Geschäftsnummer, 
                      Quartal, 
                      Land = Bestimmungsland, 
                      Wert = `Wert [CHF]`, 
                      Typ = Güterart, 
                      Signatur = `Exportkontrollnummer [EKN]`, 
                      Art = Richtung)
str(elic_2016_2)
```
### Exercise 2
We are only interested in exports (variable `Art`, value `"Ausfuhr"`), so we use
`filter()` to select the rows in the data set where this is the case. Before that
we check how many rows contain other values. `elic_2016_2$Art` gives us the variable
we are interested in and function `table()` creates a table for the frequencies of
the different values. Use table() to obtain the values of elic_2016_2$Art . 

```{r prepare-clean2, echo= FALSE}
library(ddj)
data("elic_2016_2")
#load(url("https://github.com/pbpearman/r-consortium-proposal/raw/master/material/lessons/switzerland-dual-use/data_clean/elic_2016_2.RData"))
#load(url("https://github.com/school-of-data/r-consortium-proposal/raw/master/r-package/data/elic_2016_2.RData"))
```


```{r clean2, exercise=TRUE, exercise.setup = 'prepare-clean2',cache=FALSE}

```
```{r clean2-hint}
"provide the full variable name,i.e. dataframe$variable, as an argument to table()"
```

The data set after using `filter()` should have 
`r sum(elic_2016_2$Art == "Ausfuhr")` rows.
```{r}
elic_2016_3 <- filter(.data = elic_2016_2, Art == "Ausfuhr")
str(elic_2016_3)
```
### Exercise 3
Now the collumn `Art` is not needed anymore. Let's remove it from the data set and asign the new data set to the name elic_2016_4. Use select(), specifying the elic_2016_3 data frame and the variable as -Art. Then use str() to look at the new elic_2016_4 object.
```{r rm_Art, message=FALSE, results="hide", echo=FALSE}
elic_2016_4 <- select(.data = elic_2016_3, -Art)
#str(elic_2016_4)
```
```{r prepare-clean3, echo=FALSE}
library(ddj)
data("elic_2016_3")
#load(url("https://github.com/pbpearman/r-consortium-proposal/raw/master/material/lessons/switzerland-dual-use/data_clean/elic_2016_3.RData"))
#load(url("https://github.com/school-of-data/r-consortium-proposal/raw/master/r-package/data/elic_2016_3.RData"))
```

```{r clean3, exercise=TRUE, exercise.completion=FALSE, exercise.setup = 'prepare-clean3', cache= FALSE}

```
```{r clean3-solution}
elic_2016_4 <- select(.data = elic_2016_3, -Art); str(elic_2016_4)
```


Let's see what the types of goods are in the dataset.
```{r}
unique(elic_2016_4$Typ)
```
In the output above, it looks like the variable `Typ` needs a bit of cleaning because some values are clearly variants that can be set to the simpler version. We can do that by writing a function.
```{r}
# Here are two 'vectors', the first with the original values and the second with the replacement values

typ_from <- c("Chemikalien CWÜ",
              "Dual Use Güter\r\nDual Use Güter",
              "Chemikalien CWÜ\r\nDual Use Güter",
              "Besondere militärische Güter\r\nBesondere militärische Güter",
              "Chemikalien CWÜ\r\nBesondere militärische Güter",
              "Dual Use Güter\r\nChemikalien CWÜ")
typ_to <- c("Chemikalien",
            "Dual Use Güter",
            "Chemikalien",
            "Besondere militärische Güter",
            "Chemikalien",
            "Dual Use Güter")
```
The function, below, has three arguments: `x` - the name of the column in the data set where the values are found, `from` - the vector of original values, `to` - the vector of replacement values. It contains a for-loop that updates the value of `i` to proceed through each value in 'from', identifying where in column `x` the value occurs, and replacing those values with the corresponding new value in `to`. The following lines define the function:
```{r}
change_charval <- function(x, from, to) {
  
  for(i in 1:length(from)) {
    x[x == from[i]] <- to[i]
  }
  
  return(x)
}

# Now we will call (i.e., run) the function.  We'll assign the cleaned-up values to a new variable, 'Typ', which is not in the data set.
Typ <- change_charval(x = elic_2016_4$Typ, from = typ_from, to = typ_to)

# Now we'll create a new column named 'Typ_raw' to hold the original values.
elic_2016_4$Typ_raw <- elic_2016_4$Typ

# Next we assign the cleaned-up values in 'Typ' to the data set column with the same name, effectively over-writing the original values.
elic_2016_4$Typ <- Typ
```
### Exercise 4
Use one of the R functions we used just above to check that the new values in elic_2016_4$Typ are as you expect.

```{r prepare-check1, echo=FALSE}
library(tidyverse)
library(ddj)
data("elic_2016_3")
#load(url("https://github.com/pbpearman/r-consortium-proposal/raw/master/material/lessons/switzerland-dual-use/data_clean/elic_2016_3.RData"))
#load(url("https://github.com/school-of-data/r-consortium-proposal/raw/master/r-package/data/elic_2016_3.RData"))

elic_2016_4 <- select(.data = elic_2016_3, -Art)

typ_from <- c("Chemikalien CWÜ",
              "Dual Use Güter\r\nDual Use Güter",
              "Chemikalien CWÜ\r\nDual Use Güter",
              "Besondere militärische Güter\r\nBesondere militärische Güter",
              "Chemikalien CWÜ\r\nBesondere militärische Güter",
              "Dual Use Güter\r\nChemikalien CWÜ")
typ_to <- c("Chemikalien",
            "Dual Use Güter",
            "Chemikalien",
            "Besondere militärische Güter",
            "Chemikalien",
            "Dual Use Güter")

change_charval <- function(x, from, to) {
  
  for(i in 1:length(from)) {
    x[x == from[i]] <- to[i]
  }
  
  return(x)
}
Typ <- change_charval(x = elic_2016_4$Typ, from = typ_from, to = typ_to)
elic_2016_4$Typ_raw <- elic_2016_4$Typ
elic_2016_4$Typ <- Typ
```
```{r check1, exercise=TRUE, exercise.completion=FALSE, cache= FALSE, exercise.setup = 'prepare-check1'}

```
```{r check1-hint}
"Try unique() with the correct argument value"
```



## Analyse

For this section we'll produce some nice graphics with package `ggplot2`
```{r}
library("ggplot2")
```

### A first plot
Let's try making a first plot, which shows how often Switzerland has exported
to each country. The plus sign below is particular to ggplot and means that an additional graphic element will be specified. When specified in this way, the length of the bars is determined by the number of times that a value occurs in the data. In this case, that means the number of times a country name occurs in the column `Land` in data set `elic_2016_4`.  This is determined by the ggplot aesthetic function, aes(), when specified as it is below, and a call is made to geom_bar(), which is a geom that makes bar charts.
```{r, fig.height=12, fig.width=7}

ggplot(elic_2016_4, aes(x = Land)) +  
  geom_bar() +   #this specifies the geom for a bar chart                     
  coord_flip()   #this makes the bars horizontal
```

### Order the countries by export volume

The plot above is fine if you want to check the value for a particular country, but it is hard to see trends in the data, for instance the countries that receive the most exports. Let's sort the destination countries by number of exports they receive. 

This will take a few steps. First we group the rows by country ('Land'), using `group_by()`. Then the function `summarize()` computes by country the number of exports, here 'n', actually the number of rows of entries for each country. Below, `group_by()` is nested within `summarize()`, which is allowed in R. Then `order()` identifies the row order of the data based on the number of rows 'n' associated with each country.  Then `factor()` helps us define the order in which the country levels should be plotted. `ggplot()` automatically understands that it should plot the countries in the order in which they are specified as factor levels.

```{r sort_by_number_of_exports}
# This code groups elic_2016_4 by the variable country ('Land'), then uses summarise() to find the number of rows (the entries, i.e. exports) for each country. It assigns these values to the variable 'n'. 
n_land <- summarise(group_by(elic_2016_4, Land), n = NROW(Land))
head(n_land)

# The function order() then gets the order in terms of the value of 'n', and provides an ascending index of the resulting row order
ord <- order(n_land$n)
str(ord)  # the row position of each country, in ascending order by the magnitude of n 
ordered_land <- n_land$Land[ord] # impose the sorted order and store in a variable
head(ordered_land)  #the first six values
tail(ordered_land)  #the last six values

# Now we convert the column 'Land' to the factor variable type, and set the order of the levels (i.e. values) by specifying the ordered vector of country names 'ordered_land'.  This does not change the row order of the data set, but makes the data print and plot in this order when this factor variable is used.
elic_2016_4$Land <- factor(elic_2016_4$Land, levels = ordered_land)
str(elic_2016_4$Land)

```
### Plot the countries in order of export volume
The next plot makes another horizontal bar chart in which the order of the countries is determined by the factor level we specified above.  So the bar length and the country order both depend on the number of rows associated with each country, the first because ggplot counts the number of rows, the second because elic_2016_4$Land is now an ordered factor. 

```{r, fig.height=12, fig.width=7}
ggplot(elic_2016_4, aes(x = Land)) + 
  geom_bar() +
  coord_flip()
```

### Plot countries in order of export value
The following plot shows the worth of the goods that were exported to each country. Again we want to sort the countries in a specific order, but this time the value we use to sort the data is the sum of the monitary value (the variable `Wert`) of exports to each country. So we re-order the countries based on the summed export value, and assign this order in the `levels =` specification of the call to the function factor(). By using the sorted country names in the argument `levels` argument of factor(), whenever a plot displays the levels, they are displayed in this order. 

```{r} 
wert_land <- summarise(group_by(elic_2016_4, Land), sum = sum(Wert))
# the re-ordering and assignment to a variable are here done in one line
ordered_land <- wert_land$Land[order(wert_land$sum)] 
# then the levels of the factor are assigned
elic_2016_4$Land <- factor(elic_2016_4$Land, levels = ordered_land)
```

The plot below is similar to the previous one, but by writing `stat = "identity"`, the total value of `Wert` is plotted for each country, and these are displayed in the order of magnitude of value, which we just specified above.  When you want to show bars that represent pre-calculated values, using `stat = "identity"` is the way to go.

```{r, fig.height=12, fig.width=7}
ggplot(elic_2016_4, aes(x = Land, y = Wert, fill = Typ)) + 
  geom_bar(stat = "identity") +
  coord_flip()
```

But wait, why are there countries listed, that do not seem to have gotten anything
from switzerland? Let's look at those:

```{r}
library('knitr') # kable() in this package will make nice tables

novalue  <- filter(.data = elic_2016_4, Wert == 0) # get only the lines with zero value
kable(summarize(group_by(.data=novalue,Land,Typ),n=n()),format = "html",caption = "Number Exports without Declared Value")
```
What do you think could be the potential reasons for these missing declared values?  If you are a data journalist, you may have just found something for further research.


### Countries receiving the highest value of exports
Let's look at the 10 countries that had the highest value exports in the data set.

```{r, fig.height=2, fig.width=7}
elic_2016_value <- filter(.data = elic_2016_4, Wert > 0)
elic_2016_value$Land <- droplevels(elic_2016_value$Land)
```
```{r}
elic_2016_4$Land2 <- elic_2016_4$Land
highest10 <- tail(levels(elic_2016_4$Land), 10)
```
In the following code, all the countries that aren't in the group of 10 will be pooled into "Other countries" by specifying for them just a single factor level, "Other countries".  This is done by constructing an expression that returns TRUE for the levels of elic_2016_4$Land2 (i.e., the countries) that are in highest10, using %in%, then negating it with '!', the negation operator, thus specifying the countries not in hightest10.
```{r}
head(levels(elic_2016_4$Land2) %in% highest10)
head(!levels(elic_2016_4$Land2) %in% highest10)

levels(elic_2016_4$Land2)[!(levels(elic_2016_4$Land2) %in% highest10)] <- "Other countries"
```


```{r}
wert_typ <- summarise(group_by(elic_2016_4, Typ), sum = sum(Wert))
ordered_typ <- wert_typ$Typ[order(wert_typ$sum)]
elic_2016_4$Typ <- factor(elic_2016_4$Typ, levels = rev(ordered_typ))
```
```{r, fig.width=9}
p_cntry <- ggplot(elic_2016_4, aes(x = Land2, y = Wert, fill = Typ)) + 
  geom_bar(stat = "identity") +  # so tha values in Wert are used for bar length 
  coord_flip() +  # Remember, we have to flip the plot to get horizontal bars
  xlab("Land") # There is a lot of flexibility in the labeling of the plots. Here is a start.
p_cntry
```

## Present
One cool thing about using `ggplot2` in a newsroom is that you can have different themes (color and font schemes).

```{r}
p_cntry + theme_minimal()

# It makes more sense to express the values in hundreds of millions, instead of showing the values in scientific notation, which may confuse some readers.  You can change they-axis lables using the scales package.  You can change the plot and axis titles with labs(). ggplot2 offers endless ways to customize the look of your plot.

library(scales)
p_cntry + theme_bw() +
  scale_y_continuous(labels=scales::unit_format('',1e-8)) +
  labs(title="Top Export-Länder",y="Wert in Hundert-Millionen SFR")
  

```


With the `ggthemes` package you can go even more wild and try themes that are similar to the themes of famous newsrooms...
```{r, fig.width=9}
library("ggthemes")
```

...such as the Economist,...
```{r, fig.width=9}
p_cntry + theme_economist() + scale_fill_economist()
```

...FiveThirtyEight,...
```{r, fig.width=9}
my_palette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00")
p_cntry + theme_fivethirtyeight() + scale_fill_manual(values = my_palette)
```

...or the Wall Street Journal.
```{r, fig.width=9}
p_cntry + theme_wsj() + scale_fill_wsj()
```

And if you want to really badly go back to an ungly 2003 Excel style, you 
can do that too :)
```{r}
p_cntry + theme_excel()
```

Or you take a theme you like and modify it. 
To create a School of Data themed plot, you can use the following colors and
fonts.  Gill Sans MT is used because it should work on both Windows and Mac.
```{r fig.width=9}
scodablue <- "#00AFF0" 
scodagrey <- "#333333"
scodagrey2 <- "#C9C9C9"
scodacol <- c(scodablue, "#CCCC33", "#FF358B", "#59631E", "#29D9C2", "#C9C9C9")
p_cntry + 
  scale_fill_manual(values = scodacol) +
  theme_classic(base_size = 11, base_family = "Gill Sans MT") +
  theme(axis.text = element_text(family = "Gill Sans MT Bold", colour = scodagrey),  
        panel.grid.major.x = element_line(color = scodagrey2),
        panel.grid.minor.x = element_line(color = scodagrey2), 
        line = element_line(colour = scodagrey),
        legend.position = "bottom")
```
You can learn more about ggplot [here](http://r4ds.had.co.nz/data-visualisation.html), [here](http://r4ds.had.co.nz/graphics-for-communication.html) and [here](http://ggplot2.tidyverse.org/reference/index.html).  You can easily find additional reference and instructional material. While ggplot takes some effort to learn, the ample references, on-line guides and cheat-sheets make it so that you don't have to keep everything in your head.

## Summary

We have been through the complete School of Data pipeline.  By taking the data through the pipeline, you have gained experience in some powerful and useful tools for data organization, manipulation and visualization.  You can increase your abilities in R by working through the text and questions in [R for Data Science](http://r4ds.had.co.nz/).

